---
title: "Stats 506 (Fall 2020) Final Project"
author: "Suppapat Korsurat, skorsu@umich.edu"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: readable
    toc: yes
    toc_float: true
  pdf_document:
    toc: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(knitr)
library(foreign)
library(png)
library(linguisticsdown) ## Import Picture form GitHub site.
```


```{r url, include = FALSE}
main_url <- "https://raw.githubusercontent.com/skorsu/Stats506_Final_Project/"
plot_url <- "main/Plots/"
plot_main_url <- paste0(main_url, plot_url)
```

```{r table_url, include = FALSE}
table_url <- "main/Code/(3)%20EDA%20and%20Analysis.R"
source(paste0(main_url, table_url))
```

## Introduction
There is much medical research claimed that diabetes and hypertension are 
associate with each other. Moreover, some of the studies said that diabetes is
the main factor for developing hypertension. The author curious whether this 
claim is valid for US citizens or not. Hence, this analysis's primary question 
is that "For the US people, Does diabetes affect the prevalence of 
hypertension?" In addition, the author wants to know whether this claim holds 
for each level of the aspects or not. Three aspects arouse the author's 
interests are gender, age, and obesity.

## Data and Method
The data used in this analysis is provided by the National Health and Nutrition
Examination Survey (NHANES). The author decides to use the data between the 
year 2011 to 2018.

According to the data, some respondents were collected the blood pressure data 
more than once. The author decided to use the latest record for each 
respondent. After filtering out the respondent who does not have the blood 
pressure data and the error in the data, the records with systolic less than 60
mmHG and/or diastolic should higher than 40 mmHG, the author classified the 
data into two groups: normal and hypertension. (71 words)

Apart from the blood pressure, the author will use only the respondent whose 
age is not less than 18 years old at least 18 years old as both hypertension 
and diabetes might not be detected precisely in children. Furthermore, the 
author divided the citizen into three groups: adults, middle-aged, 
and seniors. (51 words)

For the obesity status, the Body Mass Index was used to cluster the respondent
into two groups. The one who has BMI not less than 30 was considered obese; 
otherwise, he was considered no obese. (30 words)

The method used in this analysis is logistic regression. Two models were 
constructed in this analysis: the overall model and the model with the 
interested aspects. For both models, hypertension will be used as a response 
variable. In addition, the author let "no hypertension" as a reference level. 

The code used in this analysis can be found [here.](https://github.com/skorsu/Stats506_Final_Project/tree/main/Code)

## Result

The data in this analysis contained the record of 20,711 people. The figure 
below will show the number of respondents for each level of hypertension and 
diabetes.

```{r figure1, echo = FALSE, fig.cap = fig1}
include_graphics2(paste0(plot_main_url, "hypertension_diabetes.png"))
fig1 <- "Figure 1: The bar plot showing the number of respondents in 
each group and the corresponding percentage."
```

According to figure 1, 40.27% of the respondent were suffering from 
hypertension. Besides, 13.38% of the respondent had diabetes. When focusing 
on each group of diabetes, there is a difference between the group of 
hypertension. The people with diabetes had a higher chance of suffering from 
hypertension; however, the inverse trend appears in the people who did not 
have diabetes.

Apart from the plot above, the author also considered the effect of diabetes 
on hypertension for each variable, which are gender, obesity, and age group. 

```{r figure2, echo = FALSE, fig.cap = fig2}
include_graphics2(paste0(plot_main_url, "cross_all.png"))
fig2<- "Figure 2: The bar plot showing the percentage classified by 
hypertension, diabetes, and another aspect."
```

According to figure 2, most groups had a different trend from the overall 
level; those who did not have diabetes had a lower chance of suffering from 
hypertension. The author noticed that classes that were different from the 
overall trends shared the typical result; those who belong to the specific 
class have a higher/lower chance of suffering from hypertension, regardless 
of diabetes. Those classes are male, obese, all classes from the age groups. 

To be more precise about the trend and result, two logistic models were 
constructed. The first model only used diabetes as a predictor. Another model 
used all variables as predictors.

For every model, the authors let females in the gender variable, no obesity in 
the obesity variable, adults in the age group variable, and had diabetes in 
the diabetes variable be the reference levels. The result for both models 
showed in table 3.

```{r table3, echo = FALSE}
kable(output_table)
```

Figure 3: The result from both models, which includes Odd ratio, along with 
its confidence interval, and p-value.

## Conclusion
According to table 3, the author drew a conclusion for the overall level that 
the odds that one will suffer from hypertension if that person had diabetes 
is 2.04 times higher than the one who did not have diabetes.

Diabetes, nevertheless, has no effect on the prevalence of hypertension when 
considered with gender, obesity, and age as the p-value of the diabetes 
variable is higher than 0.05, the statistical significance level. 
Regardless of the suffering from diabetes, the author concluded three 
facts from the model.

  - While other factors are fixed, male respondents have a higher chance of 
  being suffering from hypertension than females by 1.61 times.
  - The ones whose BMI is greater than 30 tend to suffer from hypertension 
  than the others by 1.72 times, while other aspects are the same.
  - The older the respondent, the higher chance that one will suffer from 
  hypertension for the age groups.

To recapitulate, if considering only the relationship between hypertension and 
diabetes, one might conclude a relationship between hypertension and diabetes. 
This relationship, however, disappears when other factors kick into the 
analysis.

## Discussion
The dataset used in this analysis may have an error due to the redundant data. 
Since the author did know how NHANES do the survey, there might be possible 
that they collect the data from the same person. For example, one who is 
surveyed in 2013 might be surveyed again in 2015. If there is some information 
indicating concerning this issue, the improvement in the data cleaning process
should be performed in order to improve the quality of the result.

## Reference
- The link below is used for classifying hypertension status for each 
respondent by considering the systolic and diastolic.

  - “Understanding Blood Pressure Readings,” www.heart.org. [Online]. Available: https://www.heart.org/en/health-topics/high-blood-pressure/understanding-blood-pressure-readings.
  [Accessed: 07-Dec-2020] 

- This link is used for determining the obesity status.

  - "Obesity," www.mayoclinic.org. [Online]. Available:
  https://www.mayoclinic.org/diseases-conditions/obesity/symptoms-causes/syc-20375742. 
  [Accessed: 07-Dec-2020] 

- For the age group, the author follow this report.

  - C. D. Fryar, C. M. Hales, G. Zhang, and D. Kruszon-Moran, “Hypertension 
  Prevalence and Control Among Adults: United States, 2015–2016” [Online]. 
  Available: https://www.cdc.gov/nchs/data/databriefs/db289.pdf. 
  [Accessed: 07-Dec-2020] 

