---
title: "Stats 506 (Fall 2020) Final Project"
author: "Suppapat Korsurat, skorsu@umich.edu"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: readable
    toc: yes
    toc_float: yes
  pdf_document:
    toc: no
---

The github pages, along with code used in this analysis, can be found [here](https://github.com/skorsu/Stats506_Final_Project).

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(knitr)
library(foreign)
```


```{r url, include = FALSE}
main_url <- "https://raw.githubusercontent.com/skorsu/Stats506_Final_Project/"
branch <- "main/"
table_url <- "Code/(3)%20EDA%20and%20Analysis.R"
source(paste0(main_url, branch, table_url))
```

## Introduction
There is much medical research claimed that diabetes and hypertension are 
associate with each other. The author curious whether this claim is valid 
for US citizens or not. Hence, this analysis's primary question is that 
"For the US people, does diabetes affect the prevalence of hypertension?" 
The author also wants to know whether this claim still holds if the other 
factors, which are gender, age, and obesity, are also considered.

The primary method used in this analysis is logistic regression. The author 
will use the results from the models to conclude the effect of diabetes on 
hypertension prevalence. 

## Data and Method

The data used in this analysis is provided by the National Health and 
Nutrition Examination Survey (NHANES). The author decides to use the data 
between the year 2011 to 2018. The variables used in this analysis are 
gender, age, weight, height, diabetes, and all blood pressure records. 

Before proceeding to the analysis process, the author performed some data 
cleaning process. The first is about the blood pressure. The author decided 
to use the latest record for each respondent. The author also deleted the 
record with the systolic less than 60 mmHg and/or diastolic less than 40 mmHg 
as these might be an error from the measurement device. Besides, this analysis 
used only the respondent whose age is at least 18 years old as both 
hypertension and diabetes might not be detected precisely in children.

Apart from the data cleaning process, the author derived three additional 
variables: hypertension, obesity , and age group. Hypertension was created 
using blood pressure data. This variable classified people into two groups 
by applying the criteria that if one has systolic less than 130 mmHg and 
diastolic less than 80 mmHg, one was classified as no hypertension; 
otherwise, one was classified as hypertension. For the obesity, if one has a 
BMI greater than 30, one was considered obese; otherwise, one was considered 
not obese. For the age group, the author classified the respondent into 
three groups, adults, middle-aged, and seniors, by following the method 
proposed in this [link](https://www.cdc.gov/nchs/data/databriefs/db289.pdf).

Since the data used in the analysis is the survey data, the author needed to 
use survey weight (`WTMEC2YR`) to adjust each observation. By applying the 
method proposed in this [link](https://wwwn.cdc.gov/nchs/data/nhanes/analyticguidelines/11-16-analytic-guidelines.pdf), the survey weight is equal to `WTMEC2YR` divided by four 
since this analysis used the data from 8 years.

For the method used in this analysis, the author decided to use the test of 
independence to determine whether each variable and hypertension are 
independent or not. The survey logistic regressions were also constructed 
in this analysis by using hypertension as a predictor.

## Result

Table 1 showed the descriptive statistics for the variables used in this 
analysis after adjusting each observation by the survey weight. 

```{r table1, echo = FALSE}
cap1 <- "Table 1: Descriptive Statistics for diabetes, obesity, gender, 
and age group, along with each class's percentage in each hypertension status 
and p-value from the test of statistical independence between hypertension 
and each variable."

kable(bal_tab_all, caption = cap1)
```

According to table 1, the p-values show that all variables are not independent 
of hypertension at the significant level of 0.05. In order to be more precise 
about the relationship between hypertension and diabetes, two logistic models 
were constructed. The first model only used diabetes as a predictor. Another 
model used all variables as predictors. The result for both models showed 
in table 2.

```{r table2, echo = FALSE}
cap2 <- "Table 2: Result (Odds Ratio, along with its 95% Confidence Interval) 
of the logistic regression models, the overall model, and all-variable model. 
For the all-variable model, the percent column indicates the change in 
coefficient when constructing the simple logistic regression using only 
the particular predictor only."

kable(output_table2, caption = cap2)
```

According to table 2, the result of the all-variable model showed that 
diabetes did not have an effect on the prevalence of hypertension, which 
gainsaid with most of the traditional medical research. The author, 
hence, tried to investigate whether there was any confounding variable 
in the model or not. 

By applying the method provided in 
this [link](https://rpubs.com/josevilardy/confounding), gender is a 
confounder as the change in the coefficient is greater than 10%. Thus, the 
author decided to drop the gender variable and refit the model, shown in 
table 3.

```{r table3, echo = FALSE}
cap3 <- "Table 3: Result (Odds Ratio, along with its 95% Confidence Interval) 
of the logistic regression models without gender, the confounding variable."

kable(output_table3, caption = cap3)
```

## Conclusion
Without considering other factors, the result showed that the odds that one 
will suffer from hypertension if that person had diabetes is 2.06 times higher 
than the one who did not have diabetes. Diabetes, nevertheless, has no effect 
on the prevalence of hypertension when considered with gender, obesity, and 
age as the p-value of the diabetes variable is higher than 0.05, the 
statistical significance level. This situation occurred because one of the 
variables, gender, is a confounding variable detecting by considering the 
coefficient's percent change. After excluding gender out from the analysis, 
the result showed that the odds that one will suffer from hypertension if 
that person had diabetes is 1.14 times higher than those who did not have 
diabetes. As a result, diabetes still affects hypertension, but not as high 
as the first thought.

## Discussion
Since the data used in this analysis was surveyed by NHANES, there are some 
limitations that might affect the result. 

The first limitation is about the respondent in the data. Since the author 
used the dataset from four consecutive two-year cohorts, there might be some 
chance that some respondents were collected the data more than once. Since the
analysis regarding the relationship between two diseases does not involve the 
time-related aspects, using the data collected from each person only once 
might improve the result's accuracy. 

Another limitation is the analysis variables — this analysis used only three 
aspects: gender, obesity, and age. The actual relationship, however, should 
involve other aspects such as the stress level or the occupation. Including 
other aspects into the analysis may also improve the quality of the result.

## Reference
The reference below is the source of the datasets used in this analysis,
along with how to use the survey weight.

  - "NHANES Questionnaires, Datasets, and Related Documentation," 
  https://wwwn.cdc.gov. [Online]. Available: 
  https://wwwn.cdc.gov/nchs/nhanes/Default.aspx
  [Accessed: 29-Nov-2020]
  - "National Health and Nutrition Examination Survey:Analytic Guidelines, 
  2011-2014 and 2015-2016, " https://wwwn.cdc.gov. [Online]. Available:
  https://wwwn.cdc.gov/nchs/data/nhanes/analyticguidelines/11-16-analytic-guidelines.pdf  
  [Accessed: 14-Dec-2020]

These are the references related to creating the new variables: hypertension, 
obesity, and age group.

  - "Understanding Blood Pressure Readings," www.heart.org. [Online]. Available: https://www.heart.org/en/health-topics/high-blood-pressure/understanding-blood-pressure-readings.
  [Accessed: 07-Dec-2020] 
  - "Obesity," www.mayoclinic.org. [Online]. Available:
  https://www.mayoclinic.org/diseases-conditions/obesity/symptoms-causes/syc-20375742. 
  [Accessed: 07-Dec-2020] 
  - C. D. Fryar, C. M. Hales, G. Zhang, and D. Kruszon-Moran, "Hypertension 
  Prevalence and Control Among Adults: United States, 2015–2016" [Online]. 
  Available: https://www.cdc.gov/nchs/data/databriefs/db289.pdf. 
  [Accessed: 07-Dec-2020] 
  
Below is the reference for checking the confounding variable in the 
statistical models.

  - J. Vilardy, "Assessing Confounding" [Online]. Available: 
  https://rpubs.com/josevilardy/confounding [Accessed: 11-Dec-2020]
  

